\textbf{The statistical analysis of non-linear longitudinal data in biomedical research using generalized additive models}

### 2021-03-17 Manuscripts/Manuscript v4/Manuscript_AM_v4.tex
Need to work on the position of the basis figure and the appendix

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={The statistical analysis of non-linear longitudinal data in biomedical research using generalized additive models},
  pdfauthor={Ariel Mundo; Timothy J. Muldoon; John R. Tipton},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{subfig}
\usepackage{breqn}

@@

\hypertarget{references}{%
\section{References}\label{references}}

@@

\hypertarget{appendix-appendix}{%
\appendix}

\hypertarget{simulation-1}{%
\section{Simulation}\label{simulation-1}}

\hypertarget{compound-symmetry-and-independent-errors-in-linear-and-quadratic-responses}{%
\subsection{Compound symmetry and independent errors in linear and quadratic responses}\label{compound-symmetry-and-independent-errors-in-linear-and-quadratic-responses}}

This section simulated linear and quadratic data in the same manner as in Section \ref{simulation}. The linear simulations using Figure \ref{fig:linear-cases} show in panels A and D the simulated mean responses and individual datapoints. Panels C and G show a visual interpretation of ``correlation'' in the responses: In panel C, subjects that have a value of the random error \(\varepsilon\) either above or below the mean group response are more likely to have other observations that follow the same trajectory, thereby demonstrating correlation in the response. In panel G,because the errors are independent, there is no expectation that responses are likely to follow a similar pattern. Panels D and H show the predictions from the rm-ANOVA model.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Example with linear response}
\NormalTok{example <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(}\DataTypeTok{n_time =} \DecValTok{6}\NormalTok{, }
                    \DataTypeTok{fun_type =} \StringTok{"linear"}\NormalTok{, }
                    \DataTypeTok{error_type =} \StringTok{"correlated"}\NormalTok{) \{}

  \ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\NormalTok{(fun_type }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"linear"}\NormalTok{, }\StringTok{"quadratic"}\NormalTok{)))}
    \KeywordTok{stop}\NormalTok{(}\StringTok{'fun_type must be either "linear", or "quadratic"'}\NormalTok{)}
  \ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\NormalTok{(error_type }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"correlated"}\NormalTok{, }\StringTok{"independent"}\NormalTok{)))}
    \KeywordTok{stop}\NormalTok{(}\StringTok{'fun_type must be either "correlated", or "independent"'}\NormalTok{)}

  \KeywordTok{library}\NormalTok{(tidyverse)}
  \KeywordTok{library}\NormalTok{(mvnfast)}
  \KeywordTok{library}\NormalTok{(nlme)}
  \KeywordTok{library}\NormalTok{(ggsci)}

\NormalTok{  x <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{6}\NormalTok{, }\DataTypeTok{length.out =}\NormalTok{ n_time)}
\NormalTok{  mu <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DecValTok{0}\NormalTok{, }\KeywordTok{length}\NormalTok{(x), }\DecValTok{2}\NormalTok{)}
  \CommentTok{# linear response}
  \ControlFlowTok{if}\NormalTok{ (fun_type }\OperatorTok{==}\StringTok{ "linear"}\NormalTok{) \{}
\NormalTok{    mu[, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\OperatorTok{-}\StringTok{ }\NormalTok{(}\FloatTok{0.25}\OperatorTok{*}\NormalTok{x)}\OperatorTok{+}\DecValTok{2}  
\NormalTok{    mu[, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\FloatTok{0.25}\OperatorTok{*}\NormalTok{x}\OperatorTok{+}\DecValTok{2}
\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
    \CommentTok{# nonlinear response}

\NormalTok{    mu[, }\DecValTok{1}\NormalTok{] <-}\StringTok{  }\OperatorTok{-}\NormalTok{(}\FloatTok{0.25} \OperatorTok{*}\StringTok{ }\NormalTok{x}\OperatorTok{^}\DecValTok{2}\NormalTok{) }\FloatTok{+1.5}\OperatorTok{*}\NormalTok{x}\FloatTok{-1.25}
\NormalTok{    mu[, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\NormalTok{(}\FloatTok{0.25} \OperatorTok{*}\StringTok{ }\NormalTok{x}\OperatorTok{^}\DecValTok{2}\NormalTok{) }\FloatTok{-1.5}\OperatorTok{*}\NormalTok{x}\FloatTok{+1.25}
\NormalTok{  \}}
  \CommentTok{# matplot(mu, type = 'l')}

\NormalTok{  y <-}\StringTok{ }\KeywordTok{array}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DataTypeTok{dim =} \KeywordTok{c}\NormalTok{(}\KeywordTok{length}\NormalTok{(x), }\DecValTok{2}\NormalTok{, }\DecValTok{10}\NormalTok{))}
\NormalTok{  errors <-}\StringTok{ }\KeywordTok{array}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DataTypeTok{dim =} \KeywordTok{c}\NormalTok{(}\KeywordTok{length}\NormalTok{(x), }\DecValTok{2}\NormalTok{, }\DecValTok{10}\NormalTok{))}

  \ControlFlowTok{if}\NormalTok{ (error_type }\OperatorTok{==}\StringTok{ "independent"}\NormalTok{) \{}
    \CommentTok{## independent errors}
    \ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{) \{}
      \ControlFlowTok{for}\NormalTok{ (j }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{) \{}
\NormalTok{        errors[, i, j] <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DecValTok{6}\NormalTok{, }\DecValTok{0}\NormalTok{, }\FloatTok{0.25}\NormalTok{)}
\NormalTok{        y[, i, j] <-}\StringTok{ }\NormalTok{mu[, i] }\OperatorTok{+}\StringTok{ }\NormalTok{errors[, i, j]}
\NormalTok{      \}}
\NormalTok{    \}}
\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
    \ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{) \{     }\CommentTok{# number of treatments}
      \ControlFlowTok{for}\NormalTok{ (j }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{) \{  }\CommentTok{# number of subjects}
        \CommentTok{# compound symmetry errors}
\NormalTok{        errors[, i, j] <-}\StringTok{ }\KeywordTok{rmvn}\NormalTok{(}\DecValTok{1}\NormalTok{, }\KeywordTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{, }\KeywordTok{length}\NormalTok{(x)), }\FloatTok{0.1} \OperatorTok{*}\StringTok{ }\KeywordTok{diag}\NormalTok{(}\DecValTok{6}\NormalTok{) }\OperatorTok{+}\StringTok{ }\FloatTok{0.25} \OperatorTok{*}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{6}\NormalTok{, }\DecValTok{6}\NormalTok{))}
\NormalTok{        y[, i, j] <-}\StringTok{ }\NormalTok{mu[, i] }\OperatorTok{+}\StringTok{ }\NormalTok{errors[, i, j]}
\NormalTok{      \}}
\NormalTok{    \}}
\NormalTok{  \}    }

  \CommentTok{## subject random effects}

  \CommentTok{## visualizing the difference between independent errors and compound symmetry}
  \CommentTok{## why do we need to account for this -- overly confident inference}

  \KeywordTok{dimnames}\NormalTok{(y) <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\DataTypeTok{time =}\NormalTok{ x, }\DataTypeTok{treatment =} \DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{, }\DataTypeTok{subject =} \DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{)}
  \KeywordTok{dimnames}\NormalTok{(errors) <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\DataTypeTok{time =}\NormalTok{ x, }\DataTypeTok{treatment =} \DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{, }\DataTypeTok{subject =} \DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{)}
  \KeywordTok{dimnames}\NormalTok{(mu) <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\DataTypeTok{time =}\NormalTok{ x, }\DataTypeTok{treatment =} \DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{)}
\NormalTok{  dat <-}\StringTok{ }\KeywordTok{as.data.frame.table}\NormalTok{(y, }\DataTypeTok{responseName =} \StringTok{"y"}\NormalTok{)}
\NormalTok{  dat_errors <-}\StringTok{ }\KeywordTok{as.data.frame.table}\NormalTok{(errors, }\DataTypeTok{responseName =} \StringTok{"errors"}\NormalTok{)}
\NormalTok{  dat_mu <-}\StringTok{ }\KeywordTok{as.data.frame.table}\NormalTok{(mu, }\DataTypeTok{responseName =} \StringTok{"mu"}\NormalTok{)}
\NormalTok{  dat <-}\StringTok{ }\KeywordTok{left_join}\NormalTok{(dat, dat_errors, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"time"}\NormalTok{, }\StringTok{"treatment"}\NormalTok{, }\StringTok{"subject"}\NormalTok{))}
\NormalTok{  dat <-}\StringTok{ }\KeywordTok{left_join}\NormalTok{(dat, dat_mu, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"time"}\NormalTok{, }\StringTok{"treatment"}\NormalTok{))}
\NormalTok{  dat}\OperatorTok{$}\NormalTok{time <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(dat}\OperatorTok{$}\NormalTok{time))}
\NormalTok{  dat <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{subject =} \KeywordTok{factor}\NormalTok{(}\KeywordTok{paste}\NormalTok{(subject, treatment, }\DataTypeTok{sep =} \StringTok{"-"}\NormalTok{)))}

  \CommentTok{## repeated measures ANOVA in R}
\NormalTok{  fit_lm <-}\StringTok{ }\KeywordTok{lm}\NormalTok{(y }\OperatorTok{~}\StringTok{ }\NormalTok{time }\OperatorTok{+}\StringTok{ }\NormalTok{treatment }\OperatorTok{+}\StringTok{ }\NormalTok{time }\OperatorTok{*}\StringTok{ }\NormalTok{treatment, }\DataTypeTok{data =}\NormalTok{ dat)}
\NormalTok{  dat}\OperatorTok{$}\NormalTok{preds_lm <-}\StringTok{ }\KeywordTok{predict}\NormalTok{(fit_lm)}

\NormalTok{  fit_lme <-}\StringTok{ }\KeywordTok{lme}\NormalTok{(y }\OperatorTok{~}\StringTok{ }\NormalTok{treatment }\OperatorTok{+}\StringTok{ }\NormalTok{time }\OperatorTok{+}\StringTok{ }\NormalTok{treatment}\OperatorTok{:}\NormalTok{time,}
                 \DataTypeTok{data =}\NormalTok{ dat,}
                 \DataTypeTok{random =} \OperatorTok{~}\StringTok{ }\DecValTok{1} \OperatorTok{|}\StringTok{ }\NormalTok{subject,}
                 \DataTypeTok{correlation =} \KeywordTok{corCompSymm}\NormalTok{(}\DataTypeTok{form =} \OperatorTok{~}\StringTok{ }\DecValTok{1} \OperatorTok{|}\StringTok{ }\NormalTok{subject)}
\NormalTok{  )}

\NormalTok{  pred_dat <-}\StringTok{ }\KeywordTok{expand.grid}\NormalTok{(}
    \DataTypeTok{treatment =} \KeywordTok{factor}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{), }
    \DataTypeTok{time =} \KeywordTok{unique}\NormalTok{(dat}\OperatorTok{$}\NormalTok{time)}
\NormalTok{  )}

\NormalTok{  dat}\OperatorTok{$}\NormalTok{y_pred <-}\StringTok{ }\KeywordTok{predict}\NormalTok{(fit_lme)}

  \KeywordTok{return}\NormalTok{(}\KeywordTok{list}\NormalTok{(}
    \DataTypeTok{dat =}\NormalTok{ dat,}
    \DataTypeTok{pred_dat =}\NormalTok{ pred_dat,}
    \DataTypeTok{fit_lm =}\NormalTok{ fit_lm,}
    \DataTypeTok{fit_lme =}\NormalTok{ fit_lme}

\NormalTok{  ))}
\NormalTok{\}}

\NormalTok{plot_example <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(sim_dat) \{}
  \KeywordTok{library}\NormalTok{(patchwork)}
  \CommentTok{## Plot the simulated data}
\NormalTok{  p1 <-}\StringTok{ }\NormalTok{sim_dat}\OperatorTok{$}\NormalTok{dat }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ time, }\DataTypeTok{y =}\NormalTok{ y, }\DataTypeTok{group =}\NormalTok{ treatment, }\DataTypeTok{color =}\NormalTok{ treatment)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{show.legend=}\OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y=}\StringTok{'response'}\NormalTok{)}\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ time, }\DataTypeTok{y =}\NormalTok{ mu, }\DataTypeTok{color =}\NormalTok{ treatment),}\DataTypeTok{show.legend=}\OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{theme_classic}\NormalTok{() }\OperatorTok{+}
\StringTok{    }\CommentTok{#ggtitle("Simulated data with true response function")+}
\StringTok{    }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{plot.title =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{30}\NormalTok{, }
                                  \DataTypeTok{face =} \StringTok{"bold"}\NormalTok{),}
        \DataTypeTok{text=}\KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{30}\NormalTok{))}\OperatorTok{+}
\StringTok{    }\KeywordTok{scale_color_aaas}\NormalTok{()}

\NormalTok{  p2 <-}\StringTok{ }\NormalTok{sim_dat}\OperatorTok{$}\NormalTok{dat }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ time, }\DataTypeTok{y =}\NormalTok{ y, }\DataTypeTok{group =}\NormalTok{ subject, }\DataTypeTok{color =}\NormalTok{ treatment)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{size =} \StringTok{"Subjects"}\NormalTok{),}\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\CommentTok{# facet_wrap(~ treatment) +}
\StringTok{    }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ time, }\DataTypeTok{y =}\NormalTok{ mu, }\DataTypeTok{color =}\NormalTok{ treatment, }\DataTypeTok{size =} \StringTok{"Simulated Truth"}\NormalTok{), }\DataTypeTok{lty =} \DecValTok{1}\NormalTok{,}\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y=}\StringTok{'response'}\NormalTok{)}\OperatorTok{+}
\StringTok{    }\KeywordTok{scale_size_manual}\NormalTok{(}\DataTypeTok{name =} \StringTok{"Type"}\NormalTok{, }\DataTypeTok{values=}\KeywordTok{c}\NormalTok{(}\StringTok{"Subjects"}\NormalTok{ =}\StringTok{ }\FloatTok{0.5}\NormalTok{, }\StringTok{"Simulated Truth"}\NormalTok{ =}\StringTok{ }\DecValTok{3}\NormalTok{)) }\OperatorTok{+}
\StringTok{    }\CommentTok{#ggtitle("Simulated data\textbackslash{}nIndividual responses with population mean") +}
\StringTok{    }\KeywordTok{theme_classic}\NormalTok{()}\OperatorTok{+}
\StringTok{     }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{plot.title =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{30}\NormalTok{, }
                                \DataTypeTok{face =} \StringTok{"bold"}\NormalTok{),}
     \DataTypeTok{text=}\KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{30}\NormalTok{))}\OperatorTok{+}
\StringTok{    }\KeywordTok{scale_color_aaas}\NormalTok{()}

\NormalTok{   p3 <-}\StringTok{ }\NormalTok{sim_dat}\OperatorTok{$}\NormalTok{dat }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ time, }\DataTypeTok{y =}\NormalTok{ errors, }\DataTypeTok{group =}\NormalTok{ subject, }\DataTypeTok{color =}\NormalTok{ treatment)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{show.legend=}\OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y=}\StringTok{'errors'}\NormalTok{)}\OperatorTok{+}
\StringTok{     }\KeywordTok{theme_classic}\NormalTok{()}\OperatorTok{+}
\StringTok{    }\CommentTok{# facet_wrap(~ treatment) +}
\StringTok{    }\CommentTok{#ggtitle("Simulated errors") +}
\StringTok{     }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{plot.title =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{30}\NormalTok{, }
                                  \DataTypeTok{face =} \StringTok{"bold"}\NormalTok{),}
        \DataTypeTok{text=}\KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{30}\NormalTok{))}\OperatorTok{+}
\StringTok{    }\KeywordTok{scale_color_aaas}\NormalTok{()}

\NormalTok{  p4 <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(sim_dat}\OperatorTok{$}\NormalTok{dat, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ time, }\DataTypeTok{y =}\NormalTok{ y, }\DataTypeTok{color =}\NormalTok{ treatment)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_point}\NormalTok{()}\OperatorTok{+}\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y=}\StringTok{'response'}\NormalTok{)}\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{y =} \KeywordTok{predict}\NormalTok{(sim_dat}\OperatorTok{$}\NormalTok{fit_lme), }\DataTypeTok{group =}\NormalTok{ subject, }\DataTypeTok{size =} \StringTok{"Subjects"}\NormalTok{)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ sim_dat}\OperatorTok{$}\NormalTok{pred_dat, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{y =} \KeywordTok{predict}\NormalTok{(sim_dat}\OperatorTok{$}\NormalTok{fit_lme, }\DataTypeTok{level =} \DecValTok{0}\NormalTok{, }\DataTypeTok{newdata =}\NormalTok{ sim_dat}\OperatorTok{$}\NormalTok{pred_dat), }\DataTypeTok{size =} \StringTok{"Population"}\NormalTok{)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{scale_size_manual}\NormalTok{(}\DataTypeTok{name =} \StringTok{"Predictions"}\NormalTok{, }\DataTypeTok{values=}\KeywordTok{c}\NormalTok{(}\StringTok{"Subjects"}\NormalTok{ =}\StringTok{ }\FloatTok{0.5}\NormalTok{, }\StringTok{"Population"}\NormalTok{ =}\StringTok{ }\DecValTok{3}\NormalTok{)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{theme_classic}\NormalTok{() }\OperatorTok{+}
\StringTok{    }\CommentTok{#ggtitle("Fitted Model")+}
\StringTok{    }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{plot.title =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{30}\NormalTok{, }
                                  \DataTypeTok{face =} \StringTok{"bold"}\NormalTok{),}
        \DataTypeTok{text=}\KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{30}\NormalTok{))}\OperatorTok{+}
\StringTok{    }\KeywordTok{scale_color_aaas}\NormalTok{()}

  \KeywordTok{return}\NormalTok{((p1}\OperatorTok{+}\NormalTok{p3}\OperatorTok{+}\NormalTok{p2}\OperatorTok{+}\NormalTok{p4)}\OperatorTok{+}\KeywordTok{plot_layout}\NormalTok{(}\DataTypeTok{nrow=}\DecValTok{1}\NormalTok{)}\OperatorTok{+}\KeywordTok{plot_annotation}\NormalTok{(}\DataTypeTok{tag_levels =} \StringTok{'A'}\NormalTok{)) }

\NormalTok{\}}

\NormalTok{txt<-}\DecValTok{18}
\NormalTok{A1<-}\KeywordTok{plot_example}\NormalTok{(}\KeywordTok{example}\NormalTok{(}\DataTypeTok{fun_type =} \StringTok{"linear"}\NormalTok{, }\DataTypeTok{error_type =} \StringTok{"correlated"}\NormalTok{)) }

\NormalTok{B1<-}\KeywordTok{plot_example}\NormalTok{(}\KeywordTok{example}\NormalTok{(}\DataTypeTok{fun_type =} \StringTok{"linear"}\NormalTok{, }\DataTypeTok{error_type =} \StringTok{"independent"}\NormalTok{)) }

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={The statistical analysis of non-linear longitudinal data in biomedical research using generalized additive models},
  pdfauthor={Ariel Mundo; Timothy J. Muldoon; John R. Tipton},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{subfig}
\usepackage{breqn}

@@

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

@@
